SU($n$) spin model dynamics

### 2020-09-30 sun_phases/sun_dynamics_notes.tex
properly account for laser polarization in nuclear spin drive

\newcommand{\RR}{\mathbb{R}}
\renewcommand{\SS}{\mathbb{S}}
\newcommand{\ZZ}{\mathbb{Z}}

@@

We will generally assume unit filling, setting $L=N$, and here consider spin-orbit coupling angles $\phi\in\set{\pi/2,\pi}$\footnote{In retrospect, we note that these choices of spin-orbit coupling angles are not ideal for examining ``generic'' behavior, because they yield degenerate dispersions, as discussed in the caption of Figure \ref{fig:disps}.}.
We will also consider two types of initial states: a fully polarized state $\ket\X$ in which all spins point along the $x$ axis, and a ``back-to-back'' state $\ket\BTBX$ in which half of the spins are polarized along $+x$, and half along $-x$.
Specifically, indexing spins by their associated quasi-momentum mode $q\in[-\pi,\pi)$, the initial state $\ket\BTBX$ has spins with $q\ge0$ ($q<0$) initially polarized along $+x$ ($-x$).
In this way, the initial state $\ket\BTBX$ is correlated with $\sim\sin q$ terms of the spin Hamiltonian in \eqref{eq:spin_pair}.

@@

Figures \ref{fig:ss_spect_n2_X_p0.5}--\ref{fig:ss_spect_n2_XX_p1.0} show the time-series of $s^2(t)$, its power spectral density $\widetilde{s^2}(\omega)$, and time-series spectrum of $\bar\rho(t)$ for the case of $n=2$, and with values of $J/U$ between $10^{-2}$ and $10^1$.
In short, these figures all tell the same story: there is a ``collective'' phase when $J/U\lesssim1$, and of a ``single-particle'' phase when $J/U\gtrsim1$.
No surprises here.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Nuclear spin tomography}
\label{sec:tomography}

@@

In addition to letting atoms evolve freely under $H_{\t{spin}}$, we can turn on external control fields to address atoms' internal states.
Our primary tool here will be three lasers that off-resonantly address an excited electronic state of the atoms.
Each laser will be indexed by a different integer $m\in\set{1,0,-1}$ that determines the nuclear spin transition $\mu\to\mu+m$ accompanying the electronic $\ket\dn\to\ket\up$ transition upon absorption of a photon.
Laser $m$ will have a homogeneous drive amplitude $\Omega_m$, and a relative phase $e^{\i\phi_m}$ between adjacent lattice sites.
All lasers will be detuned by a frequency $\Delta$ away from the electronic $\ket\dn\to\ket\up$ transition.
Altogether, these lasers induce the drive Hamiltonian
\begin{align}
  H_{\t{drive}}^{\t{full}} = \sum_j \sum_{m\in\set{1,0,-1}} \Omega_m
  \p{e^{\i\phi_m j} s_{mj} \otimes\op{\up}{\dn}_j + \t{h.c.}}
  + \f{\Delta}{2} \sum_j \1_j \otimes\p{\op\up_j-\op\dn_j},
  \label{eq:drive_full}
\end{align}
where $j$ indexes a lattice site; $s_\pm$ are spin-raising ($+$) and spin-lowering ($-$) operators (containing all appropriate Clebsch-Gordan coefficients); $s_0\coloneqq s_\z$ is a spin-$z$ operator; and $\1$ is an identity operator on an $n$-level spin.

In the far-detuned limit $\Delta\gg\Omega_m$, a perturbative treatment of the excited electronic state yields an effective drive Hamiltonian that only addresses nuclear spins.
Furthermore, if the angles $\phi_m$ satisfy $\phi_m-\phi_{m'}=\p{m-m'}\phi$ for a fixed spin-orbit coupling angle $\phi$ (e.g.~with $\phi_m=m\phi$), then the gauge transformation $c_{j\mu}\to e^{\i\mu\phi j} c_{j\mu}$ makes the drive Hamiltonian spatially homogeneous.
Altogether, we arrive at a homogeneous effective drive Hamiltonian of the form
\begin{align}
  H_{\t{drive}} = \sum_j H_{\t{drive},j}^{\t{single}},
\end{align}
where the single-spin Hamiltonian $H_{\t{drive}}^{\t{single}}$ can be conveniently expressed in terms of operators $s_\x,s_\y,s_\z$ that generate rotations of an $n$-level spin (with total spin $I=\frac{n-1}{2}$) in 3D space:
\begin{align}
  H_{\t{drive}}^{\t{single}}
  = h_+ h_- s_\z + h_0 h_+ s_\x
  + h_0 h_- \p{s_\z s_\x  + s_\x s_\z}
  - h_0^2\p{s_\z^2-\v s_{\t{3D}}^2} - h_-^2 s_\x^2 - h_+^2 s_\y^2.
\end{align}
with $\v s_{\t{3D}}\equiv\p{s_\x,s_\y,s_\z}$.
Here $h_0\equiv\Omega_0/\sqrt\Delta$ and $h_\pm\equiv\p{\Omega_+\pm\Omega_-}/\sqrt{2\Delta}$ can be tuned independently through the choice of drive amplitudes $\Omega_m$; and $\v s_{\t{3D}}^2=s_\x^2+s_\y^2+s_\z^2=I\p{I+1}$ is a constant.
Table \ref{tab:drives} shows several drive Hamiltonians that can be implemented with various amplitude-matching conditions.
We do not expect these controls to be sufficient for implementing arbitrary collective SU($n$) rotations, but identifying the subgroup of SU($n$) generated by these controls could be an interesting problem for future work.
One notable capability is using pulse sequences to implement SU(2) rotations of the form $e^{-\i\v\theta\c\v s_{\t{3D}}}$ with an arbitrary rotation vector $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$.
In addition to enabling us to perform arbitrary 3D rotations, these collective rotations enable us to measure all components of the mean collective spin matrix $\bk{\v S}$, as discussed in Appendix \ref{sec:tomography}.

\begin{table}
  \centering
  \caption{Drive Hamiltonians that can be implemented with various amplitude-matching conditions.}
  \begin{tabular}{c|c|c}
    Condition & $\p{h_0,h_+,h_-}$
    & $H_{\t{drive}}^{\t{single}}$
    \\ \hline\hline
    $\Omega_+=\Omega_-=0$ & $\p{1,0,0}$ & $-s_\z^2 + \v s_{\t{3D}}^2$
    \\ \hline
    $\Omega_0=0$,  $\Omega_+=+\Omega_-$ & $\p{0,1,0}$ & $-s_\y^2$
    \\ \hline
    $\Omega_0=0$,  $\Omega_+=-\Omega_-$ & $\p{0,0,1}$ & $-s_\x^2$
    \\ \hline
    $\Omega_0=\Omega_\mp=0$ & $\p{0,1,\pm1}$
    & $\pm s_\z + s_\z^2 - \v s_{\t{3D}}^2$
    \\ \hline
    $\sqrt{2}\Omega_0=\pm\Omega_+=\pm\Omega_-$
    & $\p{1,\pm1,0}$ & $\pm s_\x + s_\x^2$
    \\ \hline
    $\sqrt{2}\Omega_0=\pm\Omega_+=\mp\Omega_-$
    & $\p{1,0,\pm1}$ & $\pm\p{s_\z s_\x+s_\x s_\z} + s_\y^2$
    \\ \hline
    $\Omega_0=\pm\Omega_+=\pm\Omega_-$ & $\p{1,\pm\sqrt{2},0}$
    & $\pm\sqrt{2} s_\x + \p{s_\x^2 - s_\y^2}$
    \\ \hline
    $\Omega_0=\pm\Omega_+=\mp\Omega_-$ & $\p{1,0,\pm\sqrt{2}}$
    & $\pm\sqrt{2}\p{s_\z s_\x + s_\x s_\z} - \p{s_\x^2-s_\y^2}$
  \end{tabular}
  \label{tab:drives}
\end{table}

@@

Here we consider a quadratic spin Hamiltonian of the general form
\begin{align}
  H = \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
  h^{\mu\nu j}_{\rho\sigma k} s_{\mu\nu j} s_{\rho\sigma k}
  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j} s_{\mu\nu j},
  \label{eq:spin}
\end{align}
where $\mu,\nu$ index orthogonal states of an $n$-level spin; $j,k\in\ZZ_N$ index one of $N$ spins; $h^{\mu\nu j}_{\rho\sigma k}$ and $\epsilon_{\mu\nu j}$ are scalars; and $s_{\mu\nu j}\coloneqq\op{\mu}{\nu}_j$ is a transition operator for spin $j$.  We can write this Hamiltonian using Schwinger bosons as
\begin{align}
  H = \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
  h^{\mu\nu j}_{\rho\sigma k}
  b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}
  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j} b_{\mu j}^\dag b_{\nu j},
\end{align}
where $b_{\mu j}$ annihilates a boson of type $\mu$ on site $j$.
The Heisenberg equations of motion for these operators are\footnote{The Hamiltonian in \eqref{eq:spin} only defines $h^{\mu\nu j}_{\rho\sigma k}$ for $j<k$.
  To simplify expressions in the remainder of this work, we therefore additionally define $h^{\rho\sigma k}_{\mu\nu j}=h^{\mu\nu j}_{\rho\sigma k}$ and $h^{\mu\nu j}_{\rho\sigma j}=0$.}
\begin{align}
  \i \d_t b_{\alpha\ell} = \sp{b_{\alpha\ell}, H}
  &= \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
  h^{\mu\nu j}_{\rho\sigma k}
  \sp{b_{\alpha\ell}, b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}}
  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j}
  \sp{b_{\alpha\ell}, b_{\mu j}^\dag b_{\nu j}} \\
  &= \sum_{\mu,\nu,\rho,\sigma,k} h^{\mu\nu\ell}_{\rho\sigma k}
  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
  b_{\rho k}^\dag b_{\sigma k}
  + \sum_{\mu,\nu} \epsilon_{\mu\nu\ell}
  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}} \\
  &= \sum_{\mu,\nu} \p{\sum_{\rho,\sigma,k}
    h^{\mu\nu\ell}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
    + \epsilon_{\mu\nu\ell}}
  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
\end{align}
where
\begin{align}
  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
  = \delta_{\mu\alpha} \delta_{\nu\alpha} b_{\alpha\ell}
  + \delta_{\mu\alpha} \p{1-\delta_{\nu\alpha}} b_{\nu\ell}
  = \delta_{\mu\alpha} b_{\nu\ell},
\end{align}
so
\begin{align}
  \i \d_t b_{\alpha\ell}
  = \sum_\nu \p{\sum_{\rho,\sigma,k}
    h^{\alpha\nu\ell}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
    + \epsilon_{\alpha\nu\ell}} b_{\nu\ell}.
\end{align}

>>>>

\newcommand{\CC}{\mathbb{C}}
\newcommand{\RR}{\mathbb{R}}
\renewcommand{\SS}{\mathbb{S}}
\newcommand{\ZZ}{\mathbb{Z}}

\newcommand{\su}{\mathfrak{su}}

@@

We will generally assume unit filling, setting $L=N$, and here consider spin-orbit coupling angles $\phi\in\set{\pi/2,\pi}$\footnote{In retrospect, we note that these choices of spin-orbit coupling angles are not ideal for examining ``generic'' behavior, because they yield degenerate dispersions, as discussed in the caption of Figure \ref{fig:disps}.}.
We will also consider two types of initial states: a fully polarized state $\ket\X$ in which all spins point along the $x$ axis, and a ``back-to-back'' state $\ket\BTBX$ in which half of the spins are polarized along $+x$, and half along $-x$.
Specifically, indexing spins by their associated quasi-momentum mode $q\in(-\pi,\pi)$, the initial state $\ket\BTBX$ has spins with $q>0$ ($q<0$) initially polarized along $+x$ ($-x$).
In this way, the initial state $\ket\BTBX$ is correlated with $\sim\sin q$ terms of the spin Hamiltonian in \eqref{eq:spin_pair}.

@@

Figures \ref{fig:ss_spect_n2_X_p0.5}--\ref{fig:ss_spect_n2_XX_p1.0} show the time-series of $s^2(t)$, its power spectral density $\widetilde{s^2}(\omega)$, and the time-series spectrum of $\bar\rho(t)$ for the case of $n=2$, with values of $J/U$ between $10^{-2}$ and $10^1$.
In short, these figures all tell the same story: there is a ``collective'' phase when $J/U\lesssim1$, and of a ``single-particle'' phase when $J/U\gtrsim1$.
No surprises here.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Spin qudit tomography}
\label{sec:tomography}

% Motivations / background:
% https://arxiv.org/pdf/1901.06282.pdf
% https://arxiv.org/abs/1911.00099
% https://arxiv.org/abs/2005.10910
% https://journals.aps.org/pra/abstract/10.1103/PhysRevA.86.012118
% https://journals.aps.org/pra/abstract/10.1103/PhysRevA.94.010301
% https://arxiv.org/pdf/2003.08668.pdf
% https://www.osapublishing.org/ol/abstract.cfm?uri=ol-44-10-2558

@@

In addition to letting atoms evolve freely under $H_{\t{spin}}$, we can turn on external control fields to address atoms' internal states.
Our primary tool here will be a multi-laser drive that off-resonantly addresses an excited electronic state of the atoms.
Specifically, all lasers will have a frequency that is detuned by $\Delta$ from an electronic $\ket\dn\to\ket\up$ transition.
Each laser will be indexed by an axis of propagation, $\v v$, with right/left-circular polarization amplitudes $\Omega_{\v v\pm} e^{\i\eta_{\v v\pm}}$ (with real $\Omega_{\v v\pm}$ and $\eta_{\v v\pm}$).
An electronic $\ket\dn\to\ket\up$ transition is thus accompanied by a nuclear spin transition $\mu_{\v v}\to\mu_{\v v}\pm1$, where $\mu_{\v v}$ is the projection of nuclear spin onto $\v v$.
Furthermore, laser $\v v$ will generally have a relative phase $e^{\i\kappa\v v\c\v\ell}$ between neighboring lattice sites, where $\kappa$ is the wavenumber of the drive lasers (in units of the lattice spacing) and $\v\ell$ is a fixed unit vector parallel to the lattice.
Altogether, these lasers induce the drive Hamiltonian
  H_{\t{drive}}^{\t{full}} = \sum_{j,\v v,\sigma}
  \Omega_{\v v\sigma} \p{e^{\i\eta_{\v v\sigma} + \i\kappa\v v\c\v\ell j}
    s_{\v v\sigma j} \otimes \op{\up}{\dn}_j + \t{h.c.}}
  + \f{\Delta}{2} \sum_j \1_j \otimes\p{\op\up_j-\op\dn_j},
where $s_{\v v\pm}$ are spin-raising/lowering operators along axis $\v v$ (containing all appropriate Clebsch-Gordan coefficients), and $\1$ is the identity operator for a single $n$-level spin (with total spin $I=\frac{n-1}{2}$).
To write out $s_{\v v\pm}$ in full, we first identify the operators $s_\x,s_\y,s_\z$ that generate rotations of an $n$-level spin in 3D space, and define $s_\pm\equiv s_\x\pm\i s_\y$.
Denoting the polar and azimuthal angles of $\v v$ respectively by $\alpha_{\v v}$ and $\beta_{\v v}$, we can use the fact that $s_\x,s_\y,s_\z$ form an $\su(2)$ algebra to expand
  s_{\v v\pm}
  &= e^{-\i\beta_{\v v}s_\z} e^{-\i\alpha_{\v v}s_\y}
  s_\pm e^{\i\alpha_{\v v}s_\y} e^{\i\beta_{\v v}s_\z} \\
  &= -\sin\alpha_{\v v}\, s_\z
  + \f12 \p{\cos\alpha_{\v v}+1} e^{\mp\i\beta_{\v v}} s_\pm
  + \f12 \p{\cos\alpha_{\v v}-1} e^{\pm\i\beta_{\v v}} s_\mp.
We now consider a particular setup involving three drive lasers: one linearly polarized laser oriented along the $x$ axis, with propagation axis $\v v_0\equiv\p{1,0,0}$ and polarization amplitudes $\Omega_{\v v_\x\sigma}e^{\i\eta_{\v_\x\sigma}}=-\Omega_0e^{\i\eta_0}/2$ (for both $\sigma\in\set{\pm1}$); two right-circularly polarized lasers pointing in opposite directions along the $z$ axis, with propagation axes $\v v_\pm\equiv\p{0,0,\pm1}$ and polarization amplitudes $\Omega_{\v v_\pm,+}e^{\i\eta_{\v_\pm,+}}=\pm\Omega_\pm e^{\i\eta_\pm}$; and a 1D lattice oriented in the $y$-$z$ plane at an angle $\theta$ from the $z$ axis, with lattice axis $\v\ell=\p{0,\sin\theta,\cos\theta}$ [\todo{make figure?}].
In this setup, the relevant spin-raising/lowering operators are
  s_{\v v_0\pm} = -s_\z \pm \i s_\y,
  s_{\v v_\pm,+} = \pm s_\pm,
and the site-dependent phases are $e^{\i\kappa\v v_m\c\v\ell j}=e^{\i m\phi j}$, where $m\in\set{0,+1,-1}$ and we define $\phi\equiv\kappa\cos\theta$ for shorthand.
Altogether, the Hamiltonian induced by this three laser drive is
  H_{\t{drive}}^{\t{3LD}} = \sum_{j,m} \Omega_m
  \p{e^{\i\eta_m+\i m\phi j} s_{mj} \otimes\op{\up}{\dn}_j + \t{h.c.}}
  + \f{\Delta}{2} \sum_j \1_j \otimes\p{\op\up_j-\op\dn_j},
where we define $s_0\equiv s_\z$ for convenience of notation.
In the far-detuned limit $\Delta\gg\Omega_m$, a perturbative treatment of the excited electronic state yields an effective drive Hamiltonian that only addresses nuclear spins.
Furthermore, the gauge transformation $c_{j\mu}\to e^{\i\mu\phi j} c_{j\mu}$ makes this effective Hamiltonian spatially homogeneous, taking the form
  H_{\t{drive}}^{\t{eff}} = \sum_j H_{\t{drive},j}^{\t{single}},
where the single-spin Hamiltonian $H_{\t{drive}}^{\t{single}}$ is given by
\begin{multline}
  \Delta \times H_{\t{drive}}^{\t{single}}
  = \p{\Omega_+^2 - \Omega_-^2} s_\z
  - \sum_{\sigma\in\set{\pm1}} \Omega_0 \Omega_\sigma
  \p{\sigma s_{\tilde\eta_\sigma,\x}
    + s_\z s_{\tilde\eta_\sigma,\x} + s_{\tilde\eta_\sigma,\x} s_\z} \\
  - \Omega_0^2 s_\z^2
  - \p{\Omega_+ + \Omega_-}^2 s_{\tilde\eta_0,\x}^2
  - \p{\Omega_+ - \Omega_-}^2 s_{\tilde\eta_0,\y}^2
  ,
  \label{eq:drive_single}
\end{multline}
with
  \tilde\eta_\pm \equiv \pm\p{\eta_0 - \eta_\pm},
  \tilde\eta_0 \equiv \f{\tilde\eta_+ - \tilde\eta_-}{2}
  = -\f{\eta_+ - \eta_-}{2},
  s_{\varphi m} = e^{-\i\varphi s_\z} s_m e^{\i\varphi s_\z}.
We do not expect these controls to be sufficient for implementing arbitrary collective SU($n$) rotations, but identifying the subgroup of SU($n$) that can be generated by \eqref{eq:drive_single} could be an interesting problem for future work.
A simple special case to consider is that of all $\eta_m=0$, in which case
  \Delta \times H_{\t{drive}}^{\t{single}}
  = \tilde\Omega_+ \tilde\Omega_- s_\z
  + \tilde\Omega_0 \tilde\Omega_- s_\x
  + \tilde\Omega_0 \tilde\Omega_+ \p{s_\z s_\x  + s_\x s_\z}
  - \tilde\Omega_0^2 s_\z^2 - \tilde\Omega_+^2 s_\x^2
  - \tilde\Omega_-^2 s_\y^2,
where we define $\tilde\Omega_0=-\Omega_0$ and $\tilde\Omega_\pm\equiv\Omega_+\pm\Omega_-$ for convenience.
Table \ref{tab:drives} shows several drive Hamiltonians that can be implemented with various amplitude-matching conditions (choices of $\tilde\Omega_m$) when all $\eta_m=0$.
One notable capability that these controls provide us is that of using pulse sequences to implement SU(2) rotations of the form $e^{-\i\v\theta\c\v s_{\t{3D}}}$, where $\v s_{\t{3D}}\equiv\p{s_\x,s_\y,s_\z}$ and $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$ is an arbitrary rotation vector.
In addition to enabling arbitrary 3D rotations, these collective rotations enable us to measure all components of the mean collective spin matrix $\bk{\v S}$, as discussed in Appendix \ref{sec:tomography}.

\begin{table}
  \centering
  \caption{Drive Hamiltonians that can be implemented with various amplitude-matching conditions when all $\eta_m=0$.
    Here $\v s_{\t{3D}}^2=s_\x^2+s_\y^2+s_\z^2=I\p{I+1}$ is a constant.}
  \begin{tabular}{c|c}
    $(\tilde\Omega_0,\tilde\Omega_+,\tilde\Omega_-)/\sqrt{\Delta}$
    & $H_{\t{drive}}^{\t{single}}$
    \\ \hline\hline
    $\p{1,0,0}$ & $-s_\z^2$
    \\ \hline
    $\p{0,1,0}$ & $-s_\x^2$
    \\ \hline
    $\p{0,0,1}$ & $-s_\y^2$
    \\ \hline
    $\p{0,1,\pm1}$ & $\pm s_\z + s_\z^2 - \v s_{\t{3D}}^2$
    \\ \hline
    $\p{1,0,\pm1}$ & $\pm s_\x + s_\x^2 - \v s_{\t{3D}}^2$
    \\ \hline
    $\p{1,\pm1,0}$
    & $\pm\p{s_\z s_\x+s_\x s_\z} + s_\y^2 - \v s_{\t{3D}}^2$
  \end{tabular}
  \label{tab:drives}
\end{table}

@@

Here we consider a quadratic spin Hamiltonian of the general form
\begin{align}
  H = \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
  h^{\mu\nu j}_{\rho\sigma k} s_{\mu\nu j} s_{\rho\sigma k}
  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j} s_{\mu\nu j},
  \label{eq:spin}
\end{align}
where $\mu,\nu$ index orthogonal states of an $n$-level spin; $j,k\in\ZZ_N$ index one of $N$ spins; $h^{\mu\nu j}_{\rho\sigma k}$ and $\epsilon_{\mu\nu j}$ are scalars; and $s_{\mu\nu j}\coloneqq\op{\mu}{\nu}_j$ is a transition operator for spin $j$.  We can write this Hamiltonian using Schwinger bosons as
\begin{align}
  H = \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
  h^{\mu\nu j}_{\rho\sigma k}
  b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}
  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j} b_{\mu j}^\dag b_{\nu j},
\end{align}
where $b_{\mu j}$ annihilates a boson of type $\mu$ on site $j$.
The Heisenberg equations of motion for these operators are\footnote{The Hamiltonian in \eqref{eq:spin} only defines $h^{\mu\nu j}_{\rho\sigma k}$ for $j<k$.
  To simplify expressions, we therefore additionally define $h^{\rho\sigma k}_{\mu\nu j}=h^{\mu\nu j}_{\rho\sigma k}$ and $h^{\mu\nu j}_{\rho\sigma j}=0$.}
\begin{align}
  \i \d_t b_{\alpha\ell} = \sp{b_{\alpha\ell}, H}
  &= \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
  h^{\mu\nu j}_{\rho\sigma k}
  \sp{b_{\alpha\ell}, b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}}
  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j}
  \sp{b_{\alpha\ell}, b_{\mu j}^\dag b_{\nu j}} \\
  &= \sum_{\mu,\nu,\rho,\sigma,k} h^{\mu\nu\ell}_{\rho\sigma k}
  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
  b_{\rho k}^\dag b_{\sigma k}
  + \sum_{\mu,\nu} \epsilon_{\mu\nu\ell}
  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}} \\
  &= \sum_{\mu,\nu} \p{\sum_{\rho,\sigma,k}
    h^{\mu\nu\ell}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
    + \epsilon_{\mu\nu\ell}}
  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
\end{align}
where
\begin{align}
  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
  = \delta_{\mu\alpha} \delta_{\nu\alpha} b_{\alpha\ell}
  + \delta_{\mu\alpha} \p{1-\delta_{\nu\alpha}} b_{\nu\ell}
  = \delta_{\mu\alpha} b_{\nu\ell},
\end{align}
so
\begin{align}
  \i \d_t b_{\alpha\ell}
  = \sum_\nu \p{\sum_{\rho,\sigma,k}
    h^{\alpha\nu\ell}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
    + \epsilon_{\alpha\nu\ell}} b_{\nu\ell}.
\end{align}
